### 2016-07-08 doc/exampleSpiePaper/fadttsterUserGuide_first.tex

\documentclass[fadttsterUserGuide_master]{subfiles}

\begin{document}
	\newpage
	\part{Example}
	\subsection{Example}
	In this example, we show how subsections and subsubsections
	are produced (there are no subsubsubsections). Note how the
	subsections are numbered.

	\subsubsection{Subexample}
	Did you note that subsubsections are not numbered? This is so in the
	\texttt{book} and \texttt{report} classes. In the \texttt{article}
	class they too have numbers. (Can you figure out why?)

	\paragraph{Note}
	Paragraphs and subparagraphs do not have numbers. And they have
	\textit{run-in} headings.

	Though named ‘‘paragraph’’ we can have several paragraphs of text
	within this.

	\subparagraph{Subnote}
	Subparagraphs have an additional indentation too.
	And they can also contain more than one paragraph of text.

\end{document}