### 2016-07-08 doc/exampleSpiePaper/fadttsterUserGuide_master.tex

\input{fadttsterUserGuide_header}

\begin{document}
	\maketitle
	\thispagestyle{empty}

	\newpage
	\tableofcontents

	\newpage	
	\subfile{fadttsterUserGuide_intro}

	\newpage	
	\subfile{fadttsterUserGuide_matlabScriptGeneration}

	\newpage	
	\subfile{fadttsterUserGuide_conclusion}

	\textit{The \LaTeX\ Companion} book \cite{latexcompanion}, the Einstein.
	journal paper \cite{einstein}, and the Donald Knuth's website \cite{knuthwebsite}. 
	The \LaTeX\ related items are \cite{latexcompanion,knuthwebsite} .
	\newpage

	\subfile{fadttsterUserGuide_appendix}

\end{document}